---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggbash

ggbash provides a bash-like REPL environment for ggplot2.

* partial match

In ggbash session, almost everything can be specified by partial match.

```{r, echo=FALSE}
library(ggplot2)
```

## Basic Usage

```{bash, eval=FALSE}
library(ggbash)
ggbash(iris) # use iris as main dataset
user@host currentDir (iris) $ p Sepal.W Sepal.L c=Sp si=Petal.W
```

```{r, eval = FALSE}
executed:
    ggplot(iris) +
    geom_point(aes(Sepal.Width,
                   Sepal.Length,
                   color=Species,
                   size=Petal.Width))
```


```{r, echo=FALSE}
ggplot(iris) +
geom_point(aes(Sepal.Width,
               Sepal.Length,
               color=Species,
               size=Petal.Width))
```


## Features

### 1. Column Index Match

```{bash, eval=FALSE}
# 'ls' checks column indices
user@host currentDir (iris) $ ls

    1: Sepal.L  (ength)
    2: Sepal.W  (idth)
    3: Petal.L  (ength)
    4: Petal.W  (idth)    	
    5: Spec     (ies)

# the same as the above 'p Sepal.W Sepal.L c=Sp si=Petal.W'
user@host currentDir (iris) $ p 2 1 c=5 si=4

```

### 2. Partial Match

In the first example, ggbash performs partial matches seven times.

* __geom names__
    + `p` matches to `geom_point`.

* __column names__
    + `Sepal.W` matches to `iris$Sepal.Width`.
  
    + `Sepal.L` matches to `iris$Sepal.Length`.
  
    + `Sp` matches to `iris$Species`.
    
    + `Petal.W` matches to `iris$Petal.Width`.

* __aesthetics names__
    + `c` matches to  `color`, which is the aesthetic of geom_point.
    + `si` matches to  `size` ('s' is ambiguous between 'shape' and 'size').

### 3. Save results

```{r, eval=FALSE}
    user@host currentDir (iris) $ cd imageDir

    user@host imageDir (iris) $ p 2 1 c=5 si=4 | png big
    saved as 'iris-Sepal.W-Sepal.L-Sp.png' (1960 x 1440)
    
    user@host imageDir (iris) $ for (i in 2:5) p 1 i | pdf --name iris-for
    saved as 'iris-for.pdf' (960 x 960)
    # TBI
```


```{r, echo = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>",
#   fig.path = "README-"
# )
```
