---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggbash: a faster way to write ggplot2

ggbash provides a bash-like REPL environment for ggplot2.

```{r, echo=FALSE}
library(ggplot2)
```

## Basic Usage

```{bash, eval=FALSE}
library(ggbash)
ggbash(iris) # start a ggbash session (using iris as main dataset)
```

```{bash, eval=FALSE}
user@host currentDir (iris) $ p Sepal.W Sepal.L c=Sp siz=Petal.W
```


```{r, eval = FALSE}
executed:
    ggplot(iris) +
    geom_point(aes(Sepal.Width,
                   Sepal.Length,
                   color=Species,
                   size=Petal.Width))
```


```{r, echo=FALSE}
ggplot(iris) +
geom_point(aes(Sepal.Width,
               Sepal.Length,
               color=Species,
               size=Petal.Width))
```


## Features

### 1. Column Index Match

```{bash, eval=FALSE}
# 'ls' checks column indices
user@host currentDir (iris) $ ls

    1: Sepal.L  (ength)
    2: Sepal.W  (idth)
    3: Petal.L  (ength)
    4: Petal.W  (idth)    	
    5: Spec     (ies)

# the same as the above 'p Sepal.W Sepal.L c=Sp si=Petal.W'
user@host currentDir (iris) $ p 2 1 c=5 si=4

# you can mix both notations
user@host currentDir (iris) $ p 2 1 c=5 si=Petal.W
```

### 2. Partial Match

In the first example, ggbash performs partial matches seven times.

* __geom names__
    + `p` matches `geom_point`.

* __column names__
    + `Sepal.W` matches `iris$Sepal.Width`.
  
    + `Sepal.L` matches `iris$Sepal.Length`.
  
    + `Sp` matches `iris$Species`.
    
    + `Petal.W` matches `iris$Petal.Width`.

* __aesthetics names__
    + `c` matches `color`, which is the aesthetic of geom_point.
    + `si` matches `size` ('s' is ambiguous within 'shape', 'size', and 'stroke').

### 3. Piping to copy/save results

```{r, eval=FALSE}
    user@host currentDir (iris) $ cd imageDir

    user@host imageDir (iris) $ p 2 1 c=5 si=4 | png big
    saved as 'iris-Sepal.W-Sepal.L-Sp.png' (big: 1960 x 1440)
    
    user@host imageDir (iris) $ for (i in 2:5) p 1 i | pdf 'iris-for'
    saved as 'iris-for.pdf' (default: 960 x 960)
    
    user@host imageDir (iris) $ p 1 2 c=spec size=4 | copy
    copied to clipboard:
    ggplot(iris) + geom_point(aes(x=Sepal.Length,
                                  y=Sepal.Width,
                                  colour=Species,
                                  size=Petal.Width))
```

## Goals

ggbash has two main goals:

1. Provide blazingly fast way to do Exploratory Data Analysis.

    + less typing by partial and fuzzy match

    + save plots with auto-generated file names

2. Make it less stressful to finalize your plots.

    + adjust colors

    + rotate axis labels

    + decide tick label intervals

    + generate line-wrapped titles

```{r, echo = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>",
#   fig.path = "README-"
# )
```
