---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggbash: A Faster Way to Write ggplot2

[![Travis-CI Build Status](https://travis-ci.org/caprice-j/ggbash.svg?branch=master)](https://travis-ci.org/caprice-j/ggbash)
[![Build status](https://ci.appveyor.com/api/projects/status/vfia7i1hfowhpqhs?svg=true)](https://ci.appveyor.com/project/caprice-j/ggbash)
[![codecov](https://codecov.io/gh/caprice-j/ggbash/branch/master/graph/badge.svg)](https://codecov.io/gh/caprice-j/ggbash)
<!-- [![Coverage Status](https://coveralls.io/repos/github/caprice-j/ggbash/badge.svg)](https://coveralls.io/github/caprice-j/ggbash) -->
[![Issue Count](https://codeclimate.com/github/caprice-j/ggbash/badges/issue_count.svg)](https://codeclimate.com/github/caprice-j/ggbash/issues)


ggbash provides a bash-like REPL environment for [ggplot2](https://github.com/tidyverse/ggplot2).

```{r, echo=FALSE}
library(ggplot2)
```

## Installation
```{r, eval=FALSE}
devtools::install_github("caprice-j/ggbash")
```

## Usage

```{bash, eval=FALSE}
library(ggbash)
ggbash(iris) # start a ggbash session (using iris as main dataset)
```

```{bash, eval=FALSE}
user@host currentDir (iris) $ p Sepal.W Sepal.L c=Sp si=Petal.W
```


```{r, eval = FALSE}
executed:
    ggplot(iris) +
    geom_point(aes(Sepal.Width,
                   Sepal.Length,
                   color=Species,
                   size=Petal.Width))
```


```{r, echo=FALSE}
ggplot(iris) +
geom_point(aes(Sepal.Width,
               Sepal.Length,
               color=Species,
               size=Petal.Width))
```

Or if you just need one figure,

```{r, eval=FALSE}
executed <- drawgg(iris, split_by_space('p Sepal.W Sepal.L c=Sp siz=Petal.W'))
copy_to_clipboard(executed$cmd)
# copied: ggplot(iris) + geom_point(aes(x=Sepal.Width, y=Sepal.Length, colour=Species, size=Petal.Width))
```

## Features

### 1. Column Index Match

```{bash, eval=FALSE}
# 'list' displays column indices
user@host currentDir (iris) $ list

    1: Sepal.L  (ength)
    2: Sepal.W  (idth)
    3: Petal.L  (ength)
    4: Petal.W  (idth)    	
    5: Spec     (ies)

# the same as the above 'p Sepal.W Sepal.L c=Sp si=Petal.W'
user@host currentDir (iris) $ p 2 1 c=5 si=4

# you can mix both notations
user@host currentDir (iris) $ p 2 1 c=5 si=Petal.W
```

Column Index Match is perhaps the fastest way to build a ggplot object.
In the above case, while the normal ggplot2 notation
(`ggplot(iris) + geom_point(aes(x=Sepal.Width, y=Sepal.Length, colour=Species, size=Petal.Width))`)
contains 90 characters (spaces not counted),
`p 2 1 c=5 si=4` is just **10 characters -- more than 80% keystroke reduction**.

With more elaborated plots, the differences become much larger.

### 2. Partial Prefix Match

In the first example (`p Sepal.W Sepal.L c=Sp si=Petal.W`),
ggbash performs partial matches seven times.

* __geom names__
    + `p` matches `geom_point`.  

* __column names__
    + `Sepal.W` matches `iris$Sepal.Width`.
  
    + `Sepal.L` matches `iris$Sepal.Length`.
  
    + `Sp` matches `iris$Species`.
    
    + `Petal.W` matches `iris$Petal.Width`.

* __aesthetics names__
    + `c` matches `color`, which is the aesthetic of geom_point.
    + `si` matches `size` ('s' is ambiguous within 'shape', 'size', and 'stroke').

Note: Approximate String Match is not supported in the current version.
    
### 3. Predefined Precedence

Even if unique identification is not possible,
`ggbash` tries to guess what is specified
instead of bluntly returning an error,
hoping to achieve least expected keystrokes.

For example, for the input of `p Sepal.W Sepal.L c=Sp s=Petal.W`,
`p` ambiguously matched four different geoms,
`geom_point`, `geom_path`, `geom_polygon`, and `geom_pointrange`.  
`ggbash` determines which geom to use
by the above predefined order of precedence 
(`point` is selected in this case).

Similarly, `s` matches three aesthetics,
`size`, `shape`, and `stroke`, preferred by this order.
Thus, `s` is interpreted as `size` aesthetic.

While it's possible to define your own precedence order
through `define_constant_list()`,
adding one or two characters might be faster in most cases.

### 4. Piping to copy/save results

```{r, eval=FALSE}
    user@host currentDir (iris) $ cd imageDir

    user@host imageDir (iris) $ p 2 1 c=5 si=4 | png big
    saved as 'iris-Sepal.W-Sepal.L-Sp.png' (big: 1960 x 1440)
    
    user@host imageDir (iris) $ for (i in 2:5) p 1 i | pdf 'iris-for'
    saved as 'iris-for.pdf' (default: 960 x 960)
    
    user@host imageDir (iris) $ p 1 2 col=Sp siz=4 | copy
    copied to clipboard:
    ggplot(iris) + geom_point(aes(x=Sepal.Length,
                                  y=Sepal.Width,
                                  colour=Species,
                                  size=Petal.Width))
```

## Goals

ggbash has two main goals:

1. Provide blazingly fast way to do Exploratory Data Analysis.

    + less typing by Column Index Match,
      Partial Prefix Match, and Predefined Precedence.

    + casualy save plots with auto-generated unique file names

2. Make it less stressful to finalize your plots.

    + adjust colors or lineweights

    + rotate axis labels

    + decide tick label intervals and limits

    + generate line-wrapped titles or legends

```{r, echo = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>",
#   fig.path = "README-"
# )
```
